---
title: 'HRS EDA: Complete Responses'
author: "Asem Berkalieva"
date: "11/28/2019"
output: pdf_document
---

## Libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data
```{r}
# all years (cleaned data)
hrs <- read.csv('../../data/hrs_complete.csv')
```

## Size
The data contains 5,322 unique individuals. 
```{r}
# total people
n_hrs <- length(unique(hrs$person_id))
n_hrs

# total people per year / drop-off
hrs %>% 
  group_by(years) %>%
  distinct(person_id) %>%
  summarise(n())
```

## Common Household

```{r}
# randomly sample one from each household
set.seed(242)
hhid_unique <- hrs %>% group_by(hhid) %>% sample_n(1) %>% select(person_id)

# filter for unique
hrs_unique <- hrs %>% filter(person_id %in% hhid_unique$person_id)
```

## New size

There now exist 3,641 subjects in our dataset such that each person comes from a unique household.
```{r}
# total people with unique household
n_hrs <- length(unique(hrs_unique$person_id))
n_hrs
```

## Education

1,774 people did not report anything for education, while 1,867 did. Of those that did, some reported two different types of education. 
```{r}
# how many did not answer: 21561 - 6241
no_edu <- hrs_unique %>% distinct(person_id, education) %>%
  filter(is.na(education))

# who answered
edu <- hrs_unique %>% distinct(person_id, education) %>%
  filter(!is.na(education)) 

# number who both did and did not answer at some point
nrow(merge(no_edu, edu, by = c("person_id", "person_id")))

# education responses
edu_response <- data.frame(cbind(type = c("no", "yes"), 
                                 responses = c(1774, 1867)))
ggplot(edu_response, aes(x = type, y = responses)) + 
  geom_bar(stat = "identity", fill = "pink2") + 
  ggtitle("Number of people who provided education information") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Response type", y = "Responses") 

# range of education
#edu_range <- edu %>% 
#  group_by(person_id) %>% 
#  summarise(education = mean(education))
```

Here is a visual of types of education responses for those who did answer:
```{r, warning=FALSE, message=FALSE}
# associate response number with edu type
edu_desc <- edu %>% mutate(edu_type = case_when(education == 0 ~ "no formal ed",
                                                 education >= 1 & education <= 11 ~ "grade school",
                                                 education == 12 ~ "high school",
                                                 education > 12 & education < 16 ~ "some college",
                                                 education == 16 ~ "college grad",
                                                 education == 17 ~ "post college",
                                                 education == 97 ~ "other",
                                                 education == 98 ~ "DK/NA",
                                                 education == 99 ~ "RF")) %>%
  mutate(edu_type_ordered = forcats::fct_relevel(edu_type, c("no formal ed",
                                               "grade school", "high school",
                                               "some college", "college grad", 
                                               "post college", "other", "DK/NA", "DF")))

# visualize types of education
ggplot(edu_desc, aes(x = edu_type_ordered)) + 
  geom_bar(fill = "pink2") + 
  ggtitle("Range of Education By Those Who Answered") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Response type", y = "Responses") 
```

## Income Range

Income responses range very heavily. The survey question asks these questions for the full past [Calendar Year].
```{r}
# range of responses
# highest amounts reported
hrs_unique %>% select(person_id, income) %>%
  arrange(-income)

# lowest amounts reported
hrs_unique %>% select(person_id, income) %>%
  arrange(income)

# overall
summary(hrs_unique$income)
ggplot(hrs_unique %>% filter(!is.na(income)), aes(x = income)) + 
  geom_histogram(col = "white", fill = "pink2", bins = 7)  + 
  ggtitle("Range of income responses") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Income") 
```

Once again, some people reported multiple incomes per year (max is 11).
```{r}
# income reported per person 
hrs_unique %>% 
  group_by(person_id) %>% 
  summarise(income_responses = sum(!is.na(income))) %>%
  arrange(-income_responses)

# example: this person answered multiple times and their answered differed
hrs_unique %>% filter(person_id == 'h090841pn032')
```

