---
title: "HRS: EDA on Completed, Subsetted Data"
author: "Asem Berkalieva"
date: "11/30/2019"
output: pdf_document
---

I now explore the final data: the one with responses from subjects for all 4 years and no more than one income response per year.

## Libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data
```{r}
# all years (cleaned data)
hrs <- read.csv('../../data/hrs_finalized.csv')
```

## Size
The data contains 3,401 unique individuals. 
```{r}
# total people
n_hrs <- length(unique(hrs$person_id))
n_hrs

# total people per year / drop-off
hrs %>% 
  group_by(years) %>%
  distinct(person_id) %>%
  summarise(n())
```

## Education

1,633 people did not report anything for education, while 1,768 did. Of those that did, some reported two different types of education. 
```{r}
# how many did not answer: 1633
no_edu <- hrs %>% distinct(person_id, education) %>%
  filter(is.na(education))

# who answered: 1768
edu <- hrs %>% distinct(person_id, education) %>%
  filter(!is.na(education)) 

# number who both did and did not answer at some point
nrow(merge(no_edu, edu, by = c("person_id", "person_id")))

# education responses
edu_response <- data.frame(cbind(type = c("no", "yes"), 
                                 responses = c(1633, 1768)))
ggplot(edu_response, aes(x = type, y = responses)) + 
  geom_bar(stat = "identity", fill = "pink2") + 
  ggtitle("Number of people who provided education information") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Response type", y = "Responses") 

# range of education
#edu_range <- edu %>% 
#  group_by(person_id) %>% 
#  summarise(education = mean(education))
```

Here is a visual of types of education responses for those who did answer:
```{r, warning=FALSE, message=FALSE}
# associate response number with edu type
edu_desc <- edu %>% mutate(edu_type = case_when(education == 0 ~ "no formal ed",
                                                 education >= 1 & education <= 11 ~ "grade school",
                                                 education == 12 ~ "high school",
                                                 education > 12 & education < 16 ~ "some college",
                                                 education == 16 ~ "college grad",
                                                 education == 17 ~ "post college",
                                                 education == 97 ~ "other",
                                                 education == 98 ~ "DK/NA",
                                                 education == 99 ~ "RF")) %>%
  mutate(edu_type_ordered = forcats::fct_relevel(edu_type, c("no formal ed",
                                               "grade school", "high school",
                                               "some college", "college grad", 
                                               "post college", "other", "DK/NA", "DF")))

# visualize types of education
ggplot(edu_desc, aes(x = edu_type_ordered)) + 
  geom_bar(fill = "pink2") + 
  ggtitle("Range of Education By Those Who Answered") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Response type", y = "Responses") 
```

## Income 
Income responses range very heavily. The survey question asks these questions for the full past [Calendar Year].
```{r}
# range of responses
# highest amounts reported
hrs %>% select(person_id, income) %>%
  arrange(-income)

# lowest amounts reported
hrs %>% select(person_id, income) %>%
  arrange(income)

# overall
summary(hrs$income)
ggplot(hrs %>% filter(!is.na(income)), aes(x = income)) + 
  geom_histogram(col = "white", fill = "pink2", bins = 7)  + 
  ggtitle("Range of income responses") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Income") 
```

## Basic summaries by time

This is a bad graph. I am trying to visualize income over time for all subjects (without looking at depression).
```{r}
# spaghetti plot of income by time per person
ggplot(hrs, aes(x = years, y = income, group = person_id)) + 
  geom_line(alpha = 0.5)
```
