---
title: "Intro to Modeling"
author: "Asem Berkalieva"
date: "11/23/2019"
output: pdf_document
---

## Libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data
```{r}
# all years (cleaned data)
hrs <- read.csv('../../data/hrs_finalized.csv')
```

## Binomial Simple GLM

The simplest form: treating every observation as independent. Not what we want, but we will build up. Our response, depression, is a binary variable, so we run a binomial regression. Our covariates are income and time (years).
```{r}
simple_bin <- glm(depression ~ income + years + income*years, 
                  data = hrs, family = "binomial")

summary(simple_bin)
```

## GEE: Independence
Accounting for repeated measures per person, we fit a binomial GEE models. First, we use 'independence' correlation structure, which reduces the model back to a simple regression, as it does not account for repeated measures within individuals:
```{r}
library(gee)
gee_ind <- gee(depression ~ income + years + income*years, person_id, 
               data = hrs, family = binomial,
               corstr = "independence")

# summary
ss_ind <- data.frame(summary(gee_ind)$coefficients)
ss_ind <- data.frame(ss_ind, pvalue = 2 * (1 - pnorm(abs(ss_ind[,5]))))
round(ss_ind, 4)
```

## GEE: Exchangeable
Now, we account for repeated measures within individuals. Results are identical to those using an independence correlation matrix.
```{r}
gee_exch <- gee(depression ~ income + years + income*years, person_id, 
                data = hrs, family = binomial,
                corstr = "exchangeable")

# summary
ss_exch <- data.frame(summary(gee_exch)$coefficients)
ss_exch <- data.frame(ss_exch, pvalue = 2 * (1 - pnorm(abs(ss_exch[,5]))))
round(ss_exch, 4)
```

## Mixed Model

```{r}
library(lme4)
glmer <- lme4::glmer(depression ~ income + years + income*years + (1|person_id), 
               data = hrs, family = binomial)

summary(glmer)
```


